Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a  Mesoscale Convective System Case During RELAMPAGO campaign

### 2022-08-22 analysis/paper/paper_ana_20181122.Rmd
ERA5 y otros comentarios menores

@@

@@

@@

  - \usepackage{natbib}

@@

  - \linenumbers 

@@

Severe weather events cause significant human and economic losses around the world. A large number of these phenomena are associated with the occurrence of deep moist convection, including tornadoes, intense wind gusts, extreme precipitation in short time periods, large hail, and lightning. 
Southern South America has one of the highest frequencies in the world of favorable conditions for high-impact meteorological events [@brooks2003] and large hail events [@cecil2012], particularly during austral spring and summer. 
This is also confirmed by observational evidence and high impact weather reports [@matsudo2015; @rasmussen2014]. Recently, the RELAMPAGO-CACTI field campaign [@nesbitt2021] has been conducted to investigate the mechanisms for convective initiation and the occurrence of high-impact weather events associated with deep convection in central Argentina.

@@

For DA methods to be successful, observing networks with sufficient temporal and spatial resolution capable of capturing mesoscale variability should be used (@gustafsson2018). In that regard, several authors investigated the impact of assimilating surface weather data (e.g. @wheatley2010, @ha2014, @chang2017, @bae2022, @banos2021, @maejima2019, and @chen2016) using different assimilation methodologies. Most of these studies reported the beneficial impacts of assimilating temperature and dew point observations upon the planetary boundary layer structure and the location and timing of precipitating systems.  @sobash2015 showed using a mesoscale DA system, that the positive impact upon convection initiation and the short range precipitation forecast is achieved if data is assimilated frequently (in the order of minutes, rather than in the order of hours).  More recently, @gasperoni2018 evaluated the impact of assimilating observations produced by private weather station networks which are not incorporated in the operational analysis. They found a positive effect of these observations upon the initiation of deep moist convection. This result is particularly important for data sparse regions such as Southern South America, where operational networks are not dense enough to capture mesoscale details.

The impact of other types of high spatial and temporal resolution observations, such as atmospheric motion vectors (AMVs), has also been investigated in the context of limited-area mesoscale DA. Many studies have focused on the impact of these observations on the prediction of tropical storms (e.g., @wu2014, @cherubini2006, and @sawada2019, and many others). Most of these studies reported an overall positive impact of the assimilation of AMVs for this type of storm. However, some works indicated mixed impacts (e.g. @sawada2019 reported an improvement in the forecast of the track of the storm but a degradation in the forecast intensity). As stated in @zhao2021 and @zhao2021a, the impact of assimilating these data on high impact weather events associated with mid-latitude deep convection over land has received relatively less attention. \citet{zhao2021, zhao2021a} assimilated GOES-16 AMVs into a storm-scale three-dimensional variational DA system during three high impact weather events. They reported positive impacts of AMVs on the characterization of the storm environment and improved short range precipitation forecasts. @otsuka2015 and @mallick2020 found a slight improvement in the short-range precipitation forecast due to the storm-scale assimilation of high frequency AMVs.

While the assimilation of radiance observations into global models is well established [@eyre2020], the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage (in the case of polar-orbiting sateññite observations), bias correction, and the relatively low model tops used for this application. @bao2015 studied the impact of assimilating microwave and infrared radiance data on temperature and humidity forecasts over the western USA and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data. More recently, @zhu2019 studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed an improvement for all variables, in particular for relative humidity at upper levels. @wang2021 studied the impact of assimilating radiances in the high-resolution Copernicus European Regional Reanalysis. They reported that satellite radiance observations had a neutral impact on the analyses of geopotential height in the lower troposphere, while a slightly negative impact on the upper troposphere and the stratosphere. They also observed similar results for 3-h forecasts initialized from the analysis but a positive impact on 12 and 24 -h forecasts. Given these mixed results, there is still room to analyze the utility of assimilating radiance observations in a limited-area DA system over land. Moreover, to the best of our knowledge, there are no studies related to the direct assimilation of radiance observations over South America.

In particular, in South America, the relatively scarcity of conventional observations such as radiosondes increases the potential impact of the observation sources previously discussed. Previous work has shown promising results of mesoscale DA in South America using some of the above mentioned data sources [e.g. @dillon2016; @dillon2021; @goncalvesdegoncalves2015]. In particular @dillon2021 assimilated high resolution surface weather station networks, GOES-16 AMVs, and satellite temperature and moisture retrievals with promising results. However, no work has been conducted yet over this region to investigate the potential contribution of the datasets discussed above. 
The main objective of this work is to contribute to the quantification and comparison of the impact of high resolution surface weather stations, AMVs, and clear-sky satellite radiances, into a mesoscale, frequently-updated ensemble-based DA system. In particular, we will focus on the impact in the context of mid-latitude mesoscale convective system events. Another particular goal of this paper is to investigate the impact of these data sources in a region where the conventional observation network is rather sparse and where the potential contributions of these observing systems is larger.

@@

Previoursly to the development of this case study, the center and north of Argentina was inmersed in a warm and humid air mass with high values of convective available potential energy (CAPE) (Figure \@ref(fig:case)a). On Nov 22, 2018 a cold front crossed the center of Argentina triggering isolated convective cells (Figure \@ref(fig:case)b) that rapidly grew upscale into an exceptionally large MCS. During that day several surface stations reported lightning, gustfronts and heavy rain. To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged with the MCS (Figure \@ref(fig:case)c). 
The MCS traveled approximately 2500 km from south to north, dissipating over Paraguay and Southern Brazil after 42 hours.

@@

campos <- ReadNetCDF(here("analysis/data/derived_data/reanalysis/geopotential.nc")) %>% 
  dcast(time + longitude + latitude ~ level) %>% 
  .[order(time, -latitude, longitude)] %>% 
  .[, ":="(msl = ReadNetCDF(here("analysis/data/derived_data/reanalysis/pressure-pw.nc"), vars = "msl", out = "vector")[[1]],
           cape = ReadNetCDF(here("analysis/data/derived_data/reanalysis/cape-pw.nc"), vars = "cape", out = "vector")[[1]],
           tcw = ReadNetCDF(here("analysis/data/derived_data/reanalysis/cape-pw.nc"), vars = "tcw", out = "vector")[[1]],
           tcwv = ReadNetCDF(here("analysis/data/derived_data/reanalysis/cape-pw.nc"), vars = "tcwv", out = "vector")[[1]],
           td = ReadNetCDF(here("analysis/data/derived_data/reanalysis/td2m.nc"), vars = "d2m", out = "vector")[[1]])]

# campos <-  ReadNetCDF(here("analysis/data/derived_data/reanalysis/cape-pw.nc"), vars = "tcw")

@@

  .[time != ymd_h(2018112312)] %>% 
  .[, ":="(espesor = (`500`-`1000`)/100, time = factor(time))] %>%
  ggplot(aes(longitude, latitude)) +
  geom_contour_fill(aes(z = cape, fill = stat(level_d)), breaks = seq(500, 4000, 500)) +
  scale_fill_distiller(super = ScaleDiscretised,
                       name = "CAPE",
                       # palette = "PRGn", direction = 1,
                       palette = "YlOrRd", direction = 1,
                       guide = guide_colorsteps(barwidth = 25, mid = 25,
                                                barheight = 0.5,
                                                title.position = "left",
                                                title.vjust = 1,
                                                frame.colour = "black")) +
  geom_mapa() +
  geom_contour2(aes(z = espesor, label = ..level..),
                size = 0.2, color = "darkred", linetype = 2,
                label_size = 2, label.placer = label_placer_n(n = 3)) +
  # label_size = 2, label.placer = label_placer_fraction(frac = 0.75)) +
  geom_contour2(aes(z = msl/100, label = ..level..),
                size = 0.4, color = "black", linetype = 1,
                label_size = 2) +
  cordillera +
  coord_sf(ylim = c(-45, -22), xlim = c(-75, -52)) +
  facet_wrap(vars(time), labeller = labeller(time = c("2018-11-22 00:00:00" = "00 UTC Nov 22",
                                                      "2018-11-22 12:00:00" = "12 UTC Nov 22",
                                                      "2018-11-23 00:00:00" = "00 UTC Nov 23"))) +
  tag_facets() +
  labs(x = NULL, y = NULL, fill = "CAPE" ) +
  theme_minimal(base_size = 8) +
  theme(legend.position = "bottom")

@@

The forecast model uses the non-hydrostatic Advanced Research version of Weather Research and Forecasting (WRF-ARW V3.9.1, @skamarock2008). 
The horizontal grid spacing is 10 km (150 x 200 grid points) in the horizontal and 37 levels in the vertical with the top of the model at 50 hPa. 
The initial and boundary conditions are provided by the Global Forecast System (GFS) analysis (0.25$^{\circ}$ horizontal grid spacing and 6-hour temporal resolution; @cisl_rda_ds084.1). 
The domain covers the area indicated in Figure \@ref(fig:dominio) to capture the development of the MCS during the simulated period.

@@

We use a 60-member ensemble where the ensemble mean at the beginning of the DA cycle is initialized using the GFS deterministic analysis with random perturbations to generate the initial ensemble perturbations. The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5$^{\circ}$ horizontal grid spacing with a smooth time evolution as in @necker2020 and @maldonado2021. In this way, we preserved the nearly hydrostatic and geostrophic equilibrium of larger scales. This method helps to prevent an underestimation of the ensemble spread [@ouaraini2015]. All experiments presented here uses the same boundary conditions from GFS, generated using the same method as for the inicial conditions to maintain proper levels of ensemble spread within the domain.

@@

To reduce the effect of spurious correlations in the estimation of error covariances, we use a horizontal localization radius of 180 km and a vertical localization radius of 0.4 (in log pressure coordinates) as in @dillon2021 for all types of observations. 
A relaxation-to-prior spread inflation [@whitaker2012] is applied with an inflation parameter $\alpha=0.9$ to mitigate the impact of sampling errors and to consider model errors not accounted for by the multi-model ensemble approach.

@@

In our experiments, only clear-sky observations are used. For microwave radiances, observations potentially contaminated by clouds are detected using the scattering and Liquid Water Path (LWP) indexes [@weston2019; @zhu2016]. For the infrared channels, cloud contaminated observations are detected using the transmittance profile calculated within the CRTM algorithms. Additionally, the GSI quality control for infrared sensors looks for observations over water with a large zenith angle (over 60°) to reject channels near the visible range that can be contaminated with reflection. It also performs an emissivity check for observations over land for both infrared and microwave radiances.

@@

@@

where $x_o$ is a assimilated observation and its error $\sigma_o$, the ensemble mean of the analysis in observational space $m$, and the standard deviation $\sigma$ of the ensemble. 
The average of $RCRV$ computed over all the realizations or analysis cycles, represents the bias of the ensemble mean with respect to the observations normalized by the estimated uncertainty:

@@

The standard deviation of the $RCRV$ or $sd RCRV$ measures the agreement of the ensemble spread and the observational error with respect to the distance between the ensemble mean and the observations, and then the systematic over- or under- dispersion of the ensemble:

\begin{equation}
  \mathrm{\mathit{sd RCRV} = \sqrt{\frac{M}{M -1}E[(\mathit{RCRV} - \mathit{mean RCRV})^2]}}
  (\#eq:eq4)
\end{equation}

where $M$ is the ensemble size. A consistent system will have no bias ($mean RCRV = 0$) and a standard deviation equal to 1 ($sd RCRV = 1$). If the ensemble has a positive bias, $mean RCRV$ will be positive, on the opposite, if the ensemble has a negative bias, $mean RCRV$ will be negative. A $sd RCRV > 1$ indicates that the difference between the observations and the ensemble mean is greater than combined uncertainty of the observations and the model, hence, the ensemble is underdispersive. On the contrary, an $sd RCRV < 1$ will indicate that the ensemble is overdispersive [@candille2007].

@@

To investigate the ability of the first-guess ensemble mean to fit the observations taking into account the uncertainties of the forecast and the observations, we calculated the $mean RCRV$ and the $sd RCRV$ for the RAD experiment. As this experiment assimilates all types of observations used in this work, it is possible to analyze the consistency of the ensemble by comparing it with each type of observation. Figure \@ref(fig:rcrv-sfc) shows the $sd RCRV$ for surface observations box-averaged to a 2.5° grid. The $sd RCRV$ for wind observations (Figure \@ref(fig:rcrv-sfc)a) is close to 1 suggesting a good agreement between the ensemble spread, the forecast error, and the observation error. For the temperature (Figure \@ref(fig:rcrv-sfc)b), the results are similar except that for some areas in the west of the domain the $sd RCRV$ can be as high as 4.5. In this region, the mean innovation was overly large due to the complex terrain leading to large differences between the model topography and the station height.

@@

Overall, these results indicate that the ensemble spread is consistent with the short-range forecast error and that systematic errors are relatively small for most of the observation types used in this work. Moreover, these results suggest the inflation parameter $\alpha = 0.9$ is adequate for the system.

@@

To investigate how changes in the PBL can modify the pre-convective environment, we compare the horizontal distribution of the low level northerly flow (for the first 7 sigma levels), precipitable water, low level temperature, and CAPE for the analysis at 00 UTC Nov 22 (after 30 assimilation cycles). At that moment the first convective cells were developing over the southern region of the domain along the cold front. Figure \@ref(fig:summary-fields)a shows the precipitable water (shaded) and the vertically averaged low-level meridional wind component (contours). We found that the moist tongue extending over the northern part of the domain is enhanced by the assimilation of denser surface observations. The moisture increase is particularly strong at the southern tip of this tongue, just ahead of the cold front where convection initiation was taking place. AWS and SATWND experiments are very similar, with values of precipitable water over 55 $kgm^{-2}$ north of 30$^{\circ}$S and a similar vertical distribution of specific humidity (not shown). RAD has lower precipitable water content than AWS and SATWND, but higher than CONV. The distribution of moisture at low levels in RAD seems to be the result of the combination of the moistening effect of assimilating AWS  -- partially compensated by the assimilation of radiance observations -- and a reduced meridional moisture transport due to the weaker northerly flow over the center of the domain compared to CONV.

@@

```{r era5, fig.cap="(ref:era5-cap)", fig.env="figure*", fig.height=7, fig.align="center"}
era <- fread("~/datosmunin3/DATA_EXP/ERA5/perfiles_ana_era5.csv")

files <- Sys.glob(here::here("/home/paola.corrales/datosmunin3/EXP//derived_data/perfiles_ana_E[2569]_new.csv"))

exp <- purrr::map(files, function(f) {
  exp <- unglue(basename(f), "perfiles_{run}_{exp}_new.csv")
  fread(f) %>% 
    .[, exp := exp[[1]][["exp"]]]

era[exp, on = c("lev", "date")] %>% 
  .[lev != 1000] %>%
  ggplot(aes(date, lev)) +
  geom_contour_fill(aes(z = T - i.T, fill = stat(level)), 
                    breaks = seq(-3, 3, 0.5)) +
  geom_contour2(aes(z = T - i.T), color = "white", size = 0.1,
                breaks = seq(-3, 3, 0.5)) +
  scale_fill_divergent_discretised(guide = guide_colorsteps(barwidth = 25,
                                                            barheight = 0.3,
                                                            title.position = "left", 
                                                            title.vjust = 1,
                                                            frame.colour = "black")) +
  scale_y_level(name = "Pressure (hPa)", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
  scale_date(ini = 20181121000000, break_bin = "12 hours") +
  facet_wrap(vars(exp), ncol = 4, labeller = labeller(exp = c(E2 = "CONV", E5 = "AWS", 
                                                              E6 = "SATWND", E9 = "RAD"))) +
  labs(fill = "Temperature (K)",
       x = NULL,
       y = "Pressure (hPa)") +
  tag_facets(position = list(x = 0.1, y = 0.96)) +
  theme_minimal(base_size = 8) +
  theme(legend.position = "bottom",
        tagger.panel.tag.text = element_text(size = 8),
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3)) +

@@

  plot_layout(ncol = 1, heights = c(1, 1, 1)) 
```

The analyzed distribution of temperature and moisture in the PBL (Figure \@ref(fig:summary-fields)b) resembles the characteristics observed in the temperature profiles (Figure \@ref(fig:TQ-diff)a-c) where AWS produces a colder PBL than CONV while the PBL in RAD is warmer than in SATWND. On average the PBL in AWS and SATWND is colder than in CONV, while RAD shows a warmer PBL than AWS due to the assimilation of radiance observations. Figure \@ref(fig:summary-fields)c shows the most unstable convective available potential energy (MCAPE, shaded) and the 0 to 6 $km$ wind shear. The values of MCAPE in CONV do not exceed 2000 $J\ Kg^{-1}$ while the rest of the experiments show maximum MCAPE over 4000 $J\ Kg^{-1}$. MCAPE in the RAD experiment is lower compared to AWS or SATWND. This is consistent with less humidity in the PBL with respect to these experiments but may be partially compensated by a slightly warmer PBL in the RAD experiment. The 0-6 km wind shear is more intense in AWS, SATWND, and RAD reaching values over 15 $m\ s^{-1}$ at the southern tip of the region with positive MCAPE values. Moreover, in this same region, these experiments show larger MCAPE values than CONV. Note that wind shear over 15 $m\ s^{-1}$ is associated with the development of more intense and organized MCSs [@chen2015] and also with conditions favorable for supercells [@markowski2010].

>>>>

  # bookdown::word_document2: default

@@

# always_allow_html: true      

@@

  - era.bib

@@

#  - \usepackage{natbib}

@@

#  - \linenumbers 

@@

Severe weather events cause significant human and economic losses around the world. A large number of these phenomena are associated with the occurrence of deep moist convection, including tornadoes, intense wind gusts, extreme precipitation in short time periods, large hail, and lightning. 
Southern South America has one of the highest frequencies in the world of favorable conditions for high-impact meteorological events [@brooks2003] and large hail events [@cecil2012], particularly during austral spring and summer. 
This is also confirmed by observational evidence and high impact weather reports [@matsudo2015; @rasmussen2014]. Recently, the Remote sensing of Electrification, Lightning, And Mesoscale/microscale Processes with Adaptive Ground Observations (RELAMPAGO) field campaign [@nesbitt2021] has been conducted to investigate the mechanisms for convective initiation and the occurrence of high-impact weather events associated with deep convection in central Argentina.

@@

For DA methods to be successful, observing networks with sufficient temporal and spatial resolution capable of capturing mesoscale variability should be used [@gustafsson2018]. Assimilating information on temperature, moisture and wind in the planetary boundary layer improves mesoscale model initialization, and several authors have reported the resultant beneficial impacts upon the planetary boundary layer structure and the location and timing of precipitating systems (e.g. @wheatley2010, @ha2014, @chang2017, @bae2022, @banos2021, @maejima2019, and @chen2016).

Particularly relevant for regional mesoscale DA systems in the region of interest is that South America is characterized by a limited number of conventional observations (i.e., radiosondes, surface weather stations) and operational networks are not dense enough to capture mesoscale details. In this context, analyzing the potential impact of non-conventional sources of observations is essential to improve mesoscale numerical weather prediction (NWP) over South America using DA. There have been only a few published efforts on regional mesoscale DA, but they have all shown promising results [e.g. @dillon2016; @dillon2021; @goncalvesdegoncalves2015]. In particular @dillon2021 assimilated high resolution surface weather station networks, GOES-16 atmospheric motion vectors (AMVs), and satellite temperature and moisture retrievals over central Argentina with positive impacts. Similarly to @gasperoni2018, @dillon2021 included private weather station networks which are not incorporated in the operational analysis. However, the impact of different observation types on the analysis quality has not been addressed.

The impact of non-conventional high spatial and temporal resolution observations, such as AMVs, has been investigated in the context of regional mesoscale DA. Many studies have focused on the impact of these observations on the prediction of tropical storms (e.g., @wu2014, @cherubini2006, and @sawada2019, and many others). Most of these studies reported an overall positive impact of the assimilation of AMVs for this type of storm. However, some works indicated mixed impacts (e.g. @sawada2019 reported an improvement in the forecast of the track of the storm but a degradation in the forecast intensity). As stated in @zhao2021 and @zhao2021a, the impact of assimilating these data on high impact weather events associated with mid-latitude deep convection over land has received relatively less attention. \citet{zhao2021, zhao2021a} assimilated GOES-16 AMVs into a storm-scale three-dimensional variational DA system during three high impact weather events. They reported positive impacts of AMVs on the characterization of the storm environment and improved short range precipitation forecasts. @otsuka2015 and @mallick2020 found a slight improvement in the short-range precipitation forecast due to the storm-scale assimilation of high frequency AMVs.

While the assimilation of radiance observations into global models is well established [@eyre2020], the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage (in the case of polar-orbiting satellite observations), bias correction, and the relatively low model tops used for this application. @bao2015 studied the impact of assimilating microwave and infrared radiance data on temperature and humidity forecasts over the western USA and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data. More recently, @zhu2019 studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed an improvement for all variables, in particular for relative humidity at upper levels. @wang2021 studied the impact of assimilating radiances in the high-resolution Copernicus European Regional Reanalysis. They reported that satellite radiance observations had a neutral impact on the analyses of geopotential height in the lower troposphere, while a slightly negative impact on the upper troposphere and the stratosphere. They also observed similar results for 3-h forecasts initialized from the analysis but a positive impact on 12 and 24 -h forecasts. Given these mixed results, there is still room to analyze the utility of assimilating radiance observations in a limited-area DA system over land. Moreover, to the best of our knowledge, there are no studies related to the direct assimilation of radiance observations over South America.

The main objective of this work is thus to contribute to the quantification and comparison of the impact of high resolution automatic weather stations, AMVs, and clear-sky satellite radiances, into a mesoscale, frequently-updated ensemble-based regional DA system. This is particularly important in the efforts to improve mesoscale numerical weather prediction (NWP) over South America where the conventional observation network is rather sparse and other sources of information could potentially fill certain gaps. In particular, we will focus on the impact in the context of a mid-latitude mesoscale convective system.

@@

Previously to the development of this case study, the center and north of Argentina was immersed in a warm and humid air mass with high values of convective available potential energy (CAPE) (Figure \@ref(fig:case)a). On Nov 22, 2018 a cold front crossed the center of Argentina triggering isolated convective cells (Figure \@ref(fig:case)b) that rapidly grew upscale into an exceptionally large MCS. During that day several surface stations reported lightning, gustfronts and heavy rain. To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged with the MCS (Figure \@ref(fig:case)c). 
The MCS traveled approximately 2500 km from south to north, dissipating over Paraguay and Southern Brazil after 42 hours.

@@

campos <- ReadNetCDF(here("analysis/data/derived_data/reanalysis/geopotential.nc")) %>% 
  dcast(time + longitude + latitude ~ level) %>% 
  .[order(time, -latitude, longitude)] %>% 
  .[, ":="(msl = ReadNetCDF(here("analysis/data/derived_data/reanalysis/pressure-pw.nc"), vars = "msl", out = "vector")[[1]],
           cape = ReadNetCDF(here("analysis/data/derived_data/reanalysis/cape-pw.nc"), vars = "cape", out = "vector")[[1]])]

@@

  .[time != ymd_h(2018112312)] %>% 
  .[, ":="(espesor = (`500`-`1000`)/100, time = factor(time))] %>%
  ggplot(aes(longitude, latitude)) +
  geom_contour_fill(aes(z = cape, fill = stat(level_d)), breaks = seq(500, 4000, 500)) +
  scale_fill_distiller(super = ScaleDiscretised,
                       name = "CAPE",
                       # palette = "PRGn", direction = 1,
                       palette = "YlOrRd", direction = 1,
                       guide = guide_colorsteps(barwidth = 25, mid = 25,
                                                barheight = 0.5,
                                                title.position = "left",
                                                title.vjust = 1,
                                                frame.colour = "black")) +
  geom_mapa() +
  geom_contour2(aes(z = espesor, label = ..level..),
                size = 0.2, color = "darkred", linetype = 2,
                label_size = 2, label.placer = label_placer_n(n = 3)) +
  geom_contour2(aes(z = msl/100, label = ..level..),
                size = 0.4, color = "black", linetype = 1,
                label_size = 2) +
  cordillera +
  coord_sf(ylim = c(-45, -22), xlim = c(-75, -52)) +
  facet_wrap(vars(time), labeller = labeller(time = c("2018-11-22 00:00:00" = "00 UTC Nov 22",
                                                      "2018-11-22 12:00:00" = "12 UTC Nov 22",
                                                      "2018-11-23 00:00:00" = "00 UTC Nov 23"))) +
  tag_facets() +
  labs(x = NULL, y = NULL, fill = "CAPE" ) +
  theme_minimal(base_size = 8) +
  theme(legend.position = "bottom")

@@

The forecast model uses the non-hydrostatic Advanced Research version of Weather Research and Forecasting (WRF-ARW V3.9.1, @skamarock2008). 
The horizontal grid spacing is 10 km (150 x 200 grid points) in the horizontal and 37 levels in the vertical with the top of the model at 50 hPa. 
The initial and boundary conditions are provided by the Global Forecast System (GFS) analysis (0.25$^{\circ}$ horizontal grid spacing and 6-hour temporal resolution; @cisl_rda_ds084.1). “In this case a single nesting approach is used since the resolution gap between the driving model and the regional model is not too large (0.25$^{\circ}$ or 25 km aproximately to 10 km). This approach is also based in recent studies which suggest that using multiple nested domains do not necessarily lead to improved precipitation forecasts in regional domains, particularly in areas of complex terrain (e.g. @liang2019, @beck2004). The domain covers the area indicated in Figure \@ref(fig:dominio) to capture the development of the MCS during the simulated period.

@@

We use a 60-member ensemble where the initial ensemble mean and the mean boundary conditions are taken from the GFS deterministic analysis. A set of 60 perturbations are  randomly generated to perturb the initial state as well as the  boundary conditions during the length of the experiment. Perturbing the boundary conditions helps to reduce the impact of errors in the driving global model and helps to keep a larger ensemble spread throughout the domain and during the length of the experiment [@ouaraini2015]. The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5$^{\circ}$ horizontal grid spacing with a smooth time evolution as in @necker2020 and @maldonado2021. In this way, we preserved the nearly hydrostatic and geostrophic equilibrium of larger scales. The random perturbations are the same across different experiments to ensure that the differences between experiments are only related to changes in the number and type of assimilated observations.

@@

To reduce the effect of spurious correlations in the estimation of error covariances, we use a horizontal localization radius of 180 km and a vertical localization radius of 0.4 (in log pressure coordinates) as in @dillon2021 for all types of observations. 
A relaxation-to-prior spread inflation [@whitaker2012] is applied with an inflation parameter $\alpha=0.9$ following @maldonado2020 to mitigate the impact of sampling errors and to consider model errors not accounted for by the multi-model ensemble approach.

@@

In our experiments, only clear-sky observations are used. For microwave radiances, observations potentially contaminated by clouds are detected using the scattering and Liquid Water Path (LWP) indexes [@weston2019; @zhu2016]. For the infrared channels, cloud contaminated observations are detected using the transmittance profile calculated within the CRTM algorithms. Moreover, GSI checks the difference between the observations and simulated brightness temperature with height to detect cloudy pixels. Additionally, the GSI quality control for infrared sensors looks for observations over water with a large zenith angle (over 60°) to reject channels near the visible range that can be contaminated with reflection. It also performs an emissivity check for observations over land for both infrared and microwave radiances.

@@

* ERA5 hourly data on pressure levels from 1959 to present [@era5pressure]. The variables of interest (air temperature, humidity and wind) were interpolated to the model grid to compare them with the analysis of each experiment.

@@

where $x_o$ is the assimilated observation and its error $\sigma_o$, the ensemble mean of the analysis in observational space $m$, and the standard deviation $\sigma$ of the ensemble.  The $RCRV$ is the ratio between the distance between the observations and the forecast and its expected standard deviation assuming the statistical independence between the forecast error (estimated from the ensemble spread) and the observation error. The average of $RCRV$ computed over all the realizations represents the bias of the ensemble mean with respect to the observations normalized by the estimated uncertainty:

@@

If the ensemble has a positive bias, $mean RCRV$ will be positive, on the opposite, if the ensemble has a negative bias, $mean RCRV$ will be negative. The standard deviation of the $RCRV$ or $sd RCRV$ is defined as:

\begin{equation}
  \mathrm{\mathit{sd RCRV} = \sqrt{\frac{1}{M -1}\sum_{i=1}^{M}(\mathit{RCRV_i} - \mathit{mean RCRV})^2}}
  (\#eq:eq4)
\end{equation}

where $M$ is the ensemble size. The $sd RCRV$ measures how large is the distance between the forecast and the observations with respect to the expected distance (given by the combination of the ensemble spread and the observation error). Assuming that the observation error is accurately estimated, an $sd RCRV > 1$ indicates that the ensemble is underdispersive (i.e. the distance between the observations and the forecasts is larger than expected) and an $sd RCRV < 1$ indicates that the ensemble is overdispersive (i.e. the distance between the observations and the forecasts is lower than expected). A consistent system will have no bias ($mean RCRV = 0$) and a standard deviation equal to 1 ($sd RCRV = 1$).

@@

To investigate the ability of the first-guess ensemble mean to fit the observations taking into account the uncertainties of the forecast and the observations, we calculated the $mean RCRV$ and the $sd RCRV$ for the RAD experiment. As this experiment assimilates all types of observations used in this work, it is possible to analyze the consistency of the ensemble by comparing it with each type of observation. Figure \@ref(fig:rcrv-sfc) shows the $sd RCRV$ for surface observations box-averaged to a 2.5° grid. The $sd RCRV$ for wind observations (Figure \@ref(fig:rcrv-sfc)a) is close to 1 suggesting a good agreement between the ensemble spread, the forecast error, and the observation error. For the temperature (Figure \@ref(fig:rcrv-sfc)b), the results are similar except that for some areas in the west of the domain the $sd RCRV$ can be as high as 4.5. These higher values of $sdRCRV$ can be associated with systematic errors arising from heigh differences between the model surface and the observations. Small scale circulations associated with the complex terrain and not well resolved by the model can also contribute to increase the distance between the forecast and the observations. These aspects are usually not captured by the ensemble spread unless a well tuned space dependent inflation scheme is used thus leading to greater sdRCRV values.

@@

Overall, these results indicate that the ensemble spread is consistent with the short-range forecast error and that systematic errors are relatively small for most of the observation types used in this work. Moreover, these results suggest the relaxation-to-prior spread inflation inflation parameter $\alpha = 0.9$ is adequate for the system.

@@

```
We also compared ensemble mean analyses to ERA5 reanalyses [@era5pressure], which suports Figures
\@ref(fig:TQ-diff) and \@ref(fig:UV-diff).  Specifically, Figure \@ref(fig:era5)a shows a warm bias in low levels (i.e. CONV is warmer than ERA5) that decreases in Figure \@ref(fig:era5)b when we assimilate the AWS observations. In the same direction, Figure \@ref(fig:TQ-diff)a shows a negative difference between AWS and CONV meaning that the AWS observations are cooling the low levels. Comparing ERA5 with RAD (Figure \@ref(fig:era5)d), there is a see a small increase in the warm bias, associated with the warming produced by the radiance observations (Figure \@ref(fig:TQ-diff)c). A similar effect can be observed for specific humidity, AWS observations parcialy corrects the dry bias present in Figure \@ref(fig:era5)e and the assimilation of radiance observations reduces this effect. The impact on the wind components is minor and only the meridional wind is included in Figure \@ref(fig:era5). In this case Figures \@ref(fig:era5)i-l show that the radiance observations are mainly responsible for the positive impact observed in the analysis. Moreover, these adjustments due to assimilating radiance and AWS observations lead to ensemble mean analyses closer to ERA5 reanalyses.

@@

```{r era5, fig.cap="(ref:era5-cap)", fig.env="figure*", fig.width=7, fig.height=6, fig.align="center"}
era <- fread(here("analysis/data/derived_data/reanalysis/perfiles_ana_era5.csv"))

era[perfiles, on = c("lev", "date")] %>% 
  .[lev != 1000] %>%
  ggplot(aes(date, lev)) +
  geom_contour_fill(aes(z = T - i.T, fill = stat(level)), 
                    breaks = seq(-3, 3, 0.5)) +
  geom_contour2(aes(z = T - i.T), color = "white", size = 0.1,
                breaks = seq(-3, 3, 0.5)) +
  scale_fill_divergent_discretised(guide = guide_colorsteps(barwidth = 25,
                                                            barheight = 0.3,
                                                            title.position = "left", 
                                                            title.vjust = 1,
                                                            frame.colour = "black")) +
  scale_y_level(name = "Pressure (hPa)", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
  scale_date(ini = 20181121000000, break_bin = "12 hours") +
  facet_wrap(vars(exp), ncol = 4, labeller = labeller(exp = c(E2 = "CONV", E5 = "AWS", 
                                                              E6 = "SATWND", E9 = "RAD"))) +
  labs(fill = "Temperature (K)",
       x = NULL,
       y = "Pressure (hPa)") +
  tag_facets(position = list(x = 0.1, y = 0.96)) +
  theme_minimal(base_size = 8) +
  theme(legend.position = "bottom",
        tagger.panel.tag.text = element_text(size = 8),
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3)) +

@@

  plot_layout(ncol = 1, heights = c(1, 1, 1)) 
```
To investigate how changes in the PBL can modify the pre-convective environment, we compare the horizontal distribution of the low level northerly flow (for the first 7 sigma levels), precipitable water, low level temperature, and CAPE for the analysis at 00 UTC Nov 22 (after 30 assimilation cycles). At that moment the first convective cells were developing over the southern region of the domain along the cold front. Figure \@ref(fig:summary-fields)a shows the precipitable water (shaded) and the vertically averaged low-level meridional wind component (contours). We found that the moist tongue extending over the northern part of the domain is enhanced by the assimilation of denser surface observations. The moisture increase is particularly strong at the southern tip of this tongue, just ahead of the cold front where convection initiation was taking place. AWS and SATWND experiments are very similar, with values of precipitable water over 55 $kgm^{-2}$ north of 30$^{\circ}$S and a similar vertical distribution of specific humidity (not shown). RAD has lower precipitable water content than AWS and SATWND, but higher than CONV. The distribution of moisture at low levels in RAD seems to be the result of the combination of the moistening effect of assimilating AWS  -- partially compensated by the assimilation of radiance observations -- and a reduced meridional moisture transport due to the weaker northerly flow over the center of the domain compared to CONV.
